---
title: "Week 2 - STA238 - Synchronous Lecture"
author: "Samantha-Jo Caetano"
date: "January 20, 2021"
output: pdf_document
---

## Law of Large Numbers (Chapter 13)

The Law of Large Numbers (LLN) states that if you take independent samples from
a well-behaved probability distribution, their sample mean converges in probability to
the true mean. Let's investigate the LLN computationally through simulations!

Load the packages we need:
```{r load-packages,warning=FALSE,message=FALSE}
library(tidyverse)
library(patchwork) # For arranging plots
options(digits = 2) # Print numbers with only 2 digits
```

Consider $X_{1},\ldots,X_{n}\overset{iid}{\sim}\text{Exponential}(\beta)$. Their
mean is $EX_{1} = 1/\beta$. If we take a sample from this distribution and 
calculate $\bar{X}_{n}$, we should get something close to $1/\beta$. If
we make $n$ larger, we should get something closer. The LLN tells us that we can
always make $n$ large enough such that we get as close as we want with as high 
a probability as we want.

To simulate from a $\text{Exponential}(\beta)$ distribution in \texttt{R}, use the
`rexp` function:

```{r sim-exp-1}
# Simulate from a exponential
# Type ?rexp to get information on the parametrization
# Parameter: rate. Rate = 1/mean.
# Always read the docs to understand what the parameters are.
# 
# Simulate with rate = 0.5, so mean = 2 and variance = ? (exercise)
rexp(1, rate = 0.5)
# The first "1" in the call gives the number of values to simulate:
rexp(10, rate = 0.5)
```

We can plot the density of the exponential sample mean for various $n$. These densities should have mean $1/\beta = 1/0.5 = 2$. 

What is the distribution of $\bar{X}$ if $X_1,..., X_n \sim Exponential(\beta)$?

Answer: $\bar{X} \sim Gamma(n, n\beta)$



```{r dens-gamma-1}
# Define a function to compute the density
# You can play around with these.
exp_samplemean_density <- function(x ,n, rate = 0.5) {
  # x: point to evaluate the density at
  # n: sample size
  
  # Just use the dgamma function
  dgamma(x,shape = n, rate = n*rate)
}

# Let's plot this:
xbar<-seq(0, 4, 0.01)
f_xbar<-exp_samplemean_density(xbar, n=5, rate=0.5)
plot(xbar, f_xbar)


# Plot it for various n
# Define a function to make the plot
plot_for_n <- function(n) {
  # Create a function with the n argument fixed
  densfun <- purrr::partial(exp_samplemean_density,n = n)
  # Plot using ggplot
  tibble(x = c(0,4)) %>%
    ggplot(aes(x = x)) +
    theme_classic() +
    stat_function(fun = densfun) +
    coord_cartesian(xlim = c(0,4),ylim = c(0,1.5)) +
    labs(title = stringr::str_c("n = ",n))
}
# Create the plots and arrange them
# You need to have the 'patchwork' package
# loaded for this to work
# library(patchwork)
(plot_for_n(1) | plot_for_n(2)) /
  (plot_for_n(4) | plot_for_n(9)) /
    (plot_for_n(16) | plot_for_n(400))
```

We can get a better look at "where the sample mean is going" by doing the following:

1. Simulate a $\text{Exponential}(0.5)$ sample of size $n$ for some really big $n$,

1. Calculate the the running average of this: for each $m = 1,\ldots,n$, compute $\bar{X}_{m} = (1/m)\sum_{i=1}^{m}X_{i}$,

1. Plot $\bar{X}_{m}$ against $m$. $\bar{X}_{m}$ should get closer and closer to $2$ as $m\to n$ (...probably).

We do this as follows:

```{r sim-exp-2}
set.seed(20210119) 

n <- 100
beta <- 0.5
exponential_experiment <- rexp(n = n,rate = beta)

# Compute the running average- a vector where the nth component is the average
# of the first n terms in exponential_experiment
runningaverage <- cumsum(exponential_experiment) / 1:length(exponential_experiment)

# Plot, remembering that the true mean is 2
tibble(x = 1:length(runningaverage),
       y = runningaverage) %>%
  ggplot(aes(x = x,y = y)) +
  theme_classic() +
  geom_point(pch = ".") +
  geom_hline(yintercept = 1/beta,colour = "red",size = .5,linetype = "dotted")

```

## Central Limit Theorem (Chapter 14) - Some Exercises


Assume that the time until a start-up company yields profits is Exponentially distributed with a mean of 2 months (i.e., rate=$\beta=0.5$). Moreover, assume (I know this is a stretch...) that start-ups (and their profits) are independent. Pretend you are a statistical consultant to some one interested in investing in some start-ups. Answer the following questions using the Central Limit Theorem and ` R` code:

1. If they invest in 30 start-ups what is the probability that the mean time until realizing profits exceeds 2.033 months (~2 months and 1 day)?
2. If they invest in 50 start-ups, what is a good range of time in which you are 95\% certain that the mean time until achieving profits will have been achieved?

```{r CLT-exp-start-ups}
## 1 P(Xbar > 2.5)
beta<-0.5
n<-30
1-pnorm(2.033, 1/beta, 1/(beta*sqrt(n)))

## 2 P(a < Xbar < b) = 0.95
beta<-0.5
n<-50

qnorm(0.025, 1/beta, 1/(beta*sqrt(n)))
qnorm(0.975, 1/beta, 1/(beta*sqrt(n)))
```

